Realizing temporal graphs from fastest travel times

### 2023-02-02 ICALP23-RealizingTemporalGraphs.tex
Changing NP-hardness for Delta (not infinity). To see old infinity versions check previous files (1 or 2 iterations back)

@@

	\begin{figure}[h]
    \centering
	\begin{tikzpicture}[xscale=2]
	%%%S_uv
	\node[vert,label=below:$u$] (v1) at (1,0) {};
	\node[vert] (v2) at (2,0) {};
	\node[vert,label=below:$w$] (v3) at (3,0) {};
	\node[vert] (v4) at (4,0) {};
	\node[vert,label=below:$v$] (v5) at (5,0) {};
	\draw (v1) -- node[timelabel] {$1$} (v2) -- node[timelabel] {$3$}  (v3) -- node[timelabel] {$5$} (v4) -- node[timelabel] {$7$} (v5);
	\draw (v1) to [out=80,in=100,distance=2cm] node[timelabel] {$9$} (v3);

@@

\begin{proof}
    Let $D$ be an input matrix for problem \deltaExact of dimension $n \times n$.
    Let us fix the vertices of the corresponding graph $G$ of $D$ as $v_1, v_2, \dots, v_n$, where vertex $v_i$ corresponds to the row and column $i$ of matrix $D$.
    This can be done in polynomial time as we need to loop through the matrix $D$ once and connect vertices $v_i, v_j$ for which $D_{i,j} = 1$. At the same time we also check if $D_{i,i} = 0$, for all $i \in [n]$.
    When $G$ is constructed we run DFS algorithm on it and check if it has no cycles.
    If at any step we encounter a problem, our algorithm stops and returns a negative answer.

@@

\section{Hardness of Exact Realization}

\todo[inline]{HM: For the following hardness I assume that we only consider strict temporal paths. (NK: Yes you are right, all paths are strict, sorry forgot to specify). \\
Furthermore, I can currently only make it work if we do not have periods ($\Delta=\infty$). I am not sure the idea also works if we have periods.}

\begin{theorem}
	\textsc{Exact Realization $(\infty)$} is \NP-hard.
\end{theorem}

\begin{proof}
	We present a polynomial-time reduction from the NP-hard problem 3-SAT~\cite{Karp1972Reducibility}. Here, we are given a formula $\phi$ in conjunctive normal form, where each clause contains exactly 3 literals (with three distinct variables). We construct an instance of \textsc{Exact Realization $(\infty)$} as follows.

 We start by describing the vertex set of the underlying graph $G$.
\begin{itemize}
\item For each variable $x$ in $\phi$, we create three variable vertices $x, x^T, x^F$.
\item For each clause $c$ in $\phi$, we create one clause vertex $c$.
\item We add one additional super vertex $v$.
\end{itemize}
Next, we describe the edge set of $G$.
\begin{itemize}
\item For each variable $x$ in $\phi$ we add the following five edges:

$\{x, x^T\}, \{x, x^F\}, \{x^T, x^F\}, \{x^T, v\}, \{x^F,v\}$.
\item For each pair of variables $x,y$ in $\phi$ with $x\neq y$ we add the following four edges

$\{x^T,y^T\},\{x^T,y^F\}, \{x^F,y^T\},\{x^F,y^F\}$.
\item For each clause $c$ in $\phi$ we add one edge for each literal. Let $x$ appear in $c$. If $X$ appears non-negated in $c$ we add edge $\{c,x^T\}$. If $x$ appears negated in $c$ we add edge $\{c, x^F\}$.
\end{itemize}
This finishes the construction of $G$.
For an example see~\cref{fig:NP-example}.

Now we specify the distances between all vertex pairs. Naturally, the distance between all pairs of adjacent vertices is one.
\begin{itemize}
    %\item For each variable $x$ in $\phi$ we specify the following distances between the non-adjacent variable vertices:

    %$d(x_1,x_2)=2$. 
    %\item For pair of variable $x,y$ in $\phi$ with $x\neq y$ we specify the following distances:

    %$d(x_1,y^T)=d(x_1,y^F)=3$, $d(x_2,y^T)=d(x_2,y^F)=2$.
    \item For each variable $x$ in $\phi$ we specify the following distances to the super vertex $v$:

    $d(x,v)=2$. %, $d(x_2,v)=3$.

    \item For each clause $c$ in $\phi$ we specify the following distances to the super vertex $v$:

    $d(c,v)=2$
    \item Let $x$ be a variable that appears in clause $c$, then  we specify the following distances:

    $d(c,x)=2$.

    If $x$ appears non-negated in $c$ we specify the following distances:

    $d(c,x^F)=2$.

    If $x$ appears negated in $c$ we specify the following distances:

    $d(c,x^T)=2$.
    \item Let $x$ be a variable that does \emph{not} appear in clause $c$, then we specify the following distances:

    $d(c,x^T)=d(c,x^F)=2$.
\end{itemize}
All distances between non-adjacent vertex pairs that are not defined above are set to $\infty$.

This finishes the construction of the \textsc{Exact Realization $(\infty)$} which can clearly be done in polynomial time. In the remainder we show that it is a yes-instance if and only if $\phi$ is satisfiable.

$(\Rightarrow)$: Assume the constructed \textsc{Exact Realization $(\infty)$} is a yes-instance. Then there exist a label $\lambda(e)$ for each edge $e\in E(G)$ such that for each vertex pair $u,w$ in the temporal graph $(G,\lambda)$ we have that a fastest temporal path between from $u$ to $w$ has exactly duration $d(u,w)$. In particular, for any two vertices $u,w$ with $d(u,w)=\infty$ we have that there does not exist a temporal path from $u$ to $w$ in $(G,\lambda)$.

We construct a satisfying assignment for $\phi$ as follows. For each variable $x$, if $\lambda(\{x,x^T\})=\lambda(\{x^T,v\})$, then we set $x$ to \texttt{true}, otherwise we set $x$ to \texttt{false}.

@@

Let $\lambda(\{x,x^T\})=t$ and $\lambda(\{x,x^F\})=t'$, for an arbitrary variable $x$. 
If both $\lambda(\{x^T,v\})\neq t+1$ and $\lambda(\{x^F,v\})\neq t'+1$, then $d(x,v)>2$, which is a contradiction. 
Thus, for every variable $x$, we have that $\lambda(\{x^T,v\})= t+1$ or $\lambda(\{x^F,v\})= t'+1$ (or both). 
In particular, this means that if $\lambda(\{x,x^F\})=\lambda(\{x^F,v\})$, then we set $x$ to \texttt{false}, since in this case $\lambda(\{x,x^T\})\neq\lambda(\{x^T,v\})$.

@@

Now assume for contradiction that the described assignment is not satisfying. Then there exists a clause $c$ that is not satisfied. Recall that we require $d(c,v)=2$. Hence, we must have a temporal path consisting of two edges from $c$ to $v$ such that the two edges have consecutive labels. By construction of $G$ there are three candidates for such a path, one for each literal of $c$. 
Assume w.l.o.g\ that $x$ appears in $c$ non-negated (the case of a negated appearance of $x$ is symmetrical) and that the temporal path realizing $d(c,v)=2$ goes through vertex $x^T$. 
Then we have that $\lambda(\{c,x^T\})=\lambda(\{x^T,v\})-1$. Furthermore, since $d(c,x)=2$ we also have that $\lambda(\{c,x^T\})=\lambda(\{x,x^T\})-1$. It follows that $\lambda(\{x,x^T\})=\lambda(\{x^T,v\})$. However, this implies that $x$ is set to \texttt{true} in the satisfying assignment and thus the clause $c$ is satisfied, a contradiction.

@@

We construct a labeling $\lambda$ as follows.
\begin{itemize}
    \item All edges incident with a clause vertex $c$ obtain label one.
    %\item For each variable $x$, we set $\lambda(\{x^T,x^F\})=1$.
    \item If variable $x$ is set to \texttt{true}, we set $\lambda(\{x^F,v\})=3$.
    \item If variable $x$ is set to \texttt{false}, we set $\lambda(\{x^T,v\})=3$.
    \item We set the labels of all other edges to two.
\end{itemize}

@@

    If $x$ is set to \texttt{true}, then there is a temporal path from $x$ to $v$ via $x^F$ such that $\lambda(\{x,x^F\})=2$ and $\lambda(\{x^F,v\})=3$. If $x$ is set to \texttt{false}, then there is a temporal path from $x$ to $v$ via $x^T$ such that $\lambda(\{x,x^T\})=2$ and $\lambda(\{x^T,v\})=3$.

@@

    Since we have a satisfying assignment there is a variable $x$ appearing in $c$ that is set to a truth-value that satisfies $c$. If $x$ appears non-negated in $c$ (and hence is set to \texttt{true}), then there is a temporal path from $c$ to $v$ through $x^T$ such that $\lambda(\{c,x^T\})=1$ and $\lambda(\{x^T,v\})=2$. If $x$ appears negated in $c$ (and hence is set to \texttt{false}), then there is a temporal path from $c$ to $v$ through $x^F$ such that $\lambda(\{c,x^F\})=1$ and $\lambda(\{x^F,v\})=2$.
    \item Let $x$ be a variable that appears in clause $c$.
    If $x$ appears non-negated in $c$ we have $d(c,x)=d(c,x_F)=2$:

    There is a temporal path from $c$ to $x$ via $x^T$ and also a temporal path from $c$ to $x^F$ via $x^T$ such that $\lambda(\{c,x^T\})=1$ and $\lambda(\{x,x^T\})=\lambda(\{x^T,x^F\})=2$.

@@

    There is a temporal path from $c$ to $x$ via $x^F$ and also a temporal path from $c$ to $x^T$ via $x^F$ such that $\lambda(\{c,x^F\})=1$ and $\lambda(\{x,x^F\})=\lambda(\{x^T,x^F\})=2$.
    \item Let $x$ be a variable that does \emph{not} appear in clause $c$, then we have $d(c,x^T)=d(c,x^F)=2$:

    Let $y$ be a variable that appears non-negated in $c$ (the case where $y$ appears negated is symmetrical). Then there is a temporal path from $c$ to $x^T$ via $y^T$ and also a temporal path from $c$ to $x^F$ via $y^T$ such that $\lambda(\{c,y^T\})=1$ and $\lambda(\{y^T,x^T\})=\lambda(\{y^T,x^F\})=2$.
\end{itemize}
Lastly, we show that all non-adjacent vertex pairs $u,w$ with $d(u,w)=\infty$ are not temporally connected in $(G,\lambda)$.
\begin{itemize}
\item For all pairs of clause vertices $c,c'$ we have $d(c,c')=\infty$: 
\todo[inline]{New matrix entries D for period Delta: d(c,c')=Delta+1 (using three edges, with labels L, L+1, L).}

@@

All edges incident with $v$ have labels two or three. All edges incident with $x$ have label two and all edges incident with $c$ have label one. We immediately have that there is no temporal path from $v$ to $c$. Furthermore, since $x$ and $v$ are non-adjacent, we also have that there is no temporal path from $v$ to $x$.
\end{itemize}
%
\begin{figure}[h]
	\noindent
	\makebox[\textwidth]{
		\centering
\begin{tikzpicture}%[xscale=2]

@@

    Given an instance $H$ of \textsc{Multicolored Clique}, we construct an instance $D$ of \deltaExactLong\ as follows. To ease the presentation, we first describe the underlying graph $G$ that is implicitly defined by the entries $D_{v,v'}=1$, that is, the pairs of vertices $v,v'$ that should be connected by a temporal path of duration one, meaning that there needs to an edge connecting the two vertices. Afterwards, we describe the remaining entries of $D$.

    We will construct $G$ using several gadgets. We first introduce an \emph{edge selection gadget} for color combination $i,j$ with $i<j$. We start with describing the vertex set of the gadget.
    \begin{itemize}
        \item A set $X_{i,j}$ of vertices $x_1, x_2, \ldots, x_m$.
        %\item A set $Y_{i,j}$ of vertices $y_1, y_2, \ldots, y_m$.
        %\item A set $Z_{i,j}$ of vertices $z_1, z_2, \ldots, z_m$.
        \item Vertex sets $U_1, U_2, \ldots, U_m$ with $U_\ell=\{u^\ell_0, u^\ell_1, u^\ell_2,\ldots, u^\ell_n, u^\ell_{n+1}, u^\ell_{n+2}, u^\ell_{2n}\}$ for all $\ell\in[m]$. Let $U_{i,j}=\bigcup_{\ell\in[n]} U_\ell$.
        \item Two special vertices $v_{i,j}^\star,v_{i,j}^{\star\star}$.
    \end{itemize}
    The gadget has the following edges.
    \begin{itemize}
        \item For all $\ell\in [m]$ we have edge $\{x_\ell,v_{i,j}^\star\}$, $\{v_{i,j}^\star,u^\ell_0\}$, and $\{u^\ell_{2n},v_{i,j}^{\star\star}\}$.

        %Furthermore, we have edges $\{y_\ell,u^\ell_1\}$ and $\{z_\ell,u^\ell_{2n}\}$.
        \item For all $\ell\in [m]$ and $\ell'\in [2n]$, we have edge $\{u^\ell_{\ell'-1},u^\ell_{\ell'}\}$.
    \end{itemize}
    This finishes the construction of the edge selection gadget $G_{i,j}$.

    For each color $i$, we introduce the following vertices. What we describe in the following will be used as a \emph{verification gadget} for color $i$.
    \begin{itemize}
        \item We have $k-1$ vertices $v^i_\ell$ for $\ell\in[k]\setminus\{i\}$ and one vertex $v^i_0$.
        \item If $1\neq i\neq k$, then we have a set $\hat{U}_i$ of $n+3$ additional vertices $\hat{u}^i_1,\hat{u}^i_2,\ldots,\hat{u}^i_{n+3}$.
    \end{itemize}
    We add the following edges. 
    %Assume $1\neq i \neq k$. We describe the two cases $i=1$ and $i=k$ afterwards. 
    %If $i\neq 1$, let $e_\ell\in F_{1,i}$ let $w^i_{\ell'}\in W_i$ be incident with $e_\ell$. Then we add edge $\{v^i_0,u^\ell_{\ell'}\}$ between $v^i_0$ and the vertex $u^\ell_{\ell'}$ of the edge selection gadget of color combination $1,i$.

    Let $1\le j<i$ (skip if $i=1$), let $e_\ell^{j,i}\in F_{j,i}$, and let $w^i_{\ell'}\in W_i$ be incident with $e_\ell^{j,i}$. Then we add edge $\{v_{j-1}^i,u^\ell_{\ell'-1}\}$ between $v^i_{j-1}$ and the vertex $u^\ell_{\ell'-1}$ of the edge selection gadget of color combination $j,i$.
    Furthermore, we add edge $\{v_{j}^i,u^\ell_{\ell'}\}$ between $v^i_j$ and the vertex $u^\ell_{\ell'}$ of the edge selection gadget of color combination $j,i$.

    If $1\neq i\neq k$, then we add edge $\{v^i_{i-1},\hat{u}^i_1\}$ and for all $\ell\in[n+2]$ we add edge $\{\hat{u}^i_\ell,\hat{u}^i_{\ell+1}\}$. 
    Let $e_\ell^{i,i+1}\in F_{i,i+1}$ and let $w^i_{\ell'}\in W_i$ be incident with $e_\ell^{i,i+1}$. If $i\neq 1\neq k$, then we add edge $\{\hat{u}^i_{n+3},u^\ell_{n+\ell'-1}\}$ between $\hat{u}^i_{n+3}$ and the vertex $u^\ell_{n+\ell'-1}$ of the edge selection gadget of color combination $i,i+1$. If $i=1$, then we add edge $\{v^i_0,u^\ell_{n+\ell'-1}\}$ between $v^i_0$ and the vertex $u^\ell_{n+\ell'-1}$ of the edge selection gadget of color combination $i,i+1$.

    Let $i<j\le k$ (skip if $i=k$), let $e_\ell^{i,j}\in F_{i,j}$, and let $w^i_{\ell'}\in W_i$ be incident with $e_\ell^{i,j}$. Then we add edge $\{v_{j-1}^i,u^\ell_{n+\ell'-1}\}$ between $v^i_{j-1}$ and the vertex $u^\ell_{n+\ell'-1}$ of the edge selection gadget of color combination $i,j$.
    Furthermore, we add edge $\{v_{j}^i,u^\ell_{n+\ell'}\}$ between $v^i_j$ and the vertex $u^\ell_{n+\ell'}$ of the edge selection gadget of color combination $i,j$.

    Next, we describe \emph{connector gadgets}. Intuitively, these gadgets will be used to connect many vertex pairs by short paths, which will make arguing about possible labelings in yes-instances much easier. Connector gadgets consist of six vertices $\hat{v}_0,\hat{v}_0',\hat{v}_1,\hat{v}_2,\hat{v}_3,\hat{v}_3'$. 
    %Each connector gadget is associated with four sets $A,A',B,B'$ containing vertices of other gadgets. 
    Each connector gadget is associated with two sets $A,B$ with $B\subseteq A$ containing vertices of other gadgets. 
    Let $V^\star$ denote the set of all vertices from all edge selection gadgets and all validation gadgets.
    The sets $A$ and $B$ will only play a role when defining the distance matrix $D$ later. Informally speaking, vertices in $A$ should reach all vertices in $V^\star$ quickly through the gadget, except the ones in $B$.
    We have the following edges. 
    \begin{itemize}
        \item Edges $\{\hat{v}_0,\hat{v}_1\},\{\hat{v}_0',\hat{v}_1\},\{\hat{v}_1,\hat{v}_2\},\{\hat{v}_2,\hat{v}_3\},\{\hat{v}_2,\hat{v}_3'\}$.
        \item An edge between $\hat{v}_1$ and each vertex in $V^\star$.
        \item An edge between $\hat{v}_2$ and each vertex in $V^\star$.
    \end{itemize}
    We add two connector gadgets for each edge selection gadget and one connector gadget for each validation gadget.

>>>>

SKELETON:
\begin{itemize}
    \item graph realization problems in general, well studied and motivated for different types of graph attributes (among the most natural ones are distances and degree sequences), mention the two perspectives: verification and network design. 
    \item Some observations and references about distance realization known results on static graphs.
    \item In modern network applications, network structure changes over time
    \item In this paper we focus on distance attributes (one of the most fundamental graph attributes)
    \item What is distances in temporal graphs? Fastest, etc, periodic
    \item Our contributions
    \item Related work on temporal network design etc
\end{itemize}

@@

	\begin{figure}[htb]
    \centering
	\begin{tikzpicture}[xscale=2]
	%%%S_uv
	\node[vert,label=below:$u$] (v1) at (1,0) {};
	\node[vert] (v2) at (2,0) {};
	\node[vert,label=below:$w$] (v3) at (3,0) {};
	\node[vert] (v4) at (4,0) {};
	\node[vert,label=below:$v$] (v5) at (5,0) {};
	\draw (v1) -- node[timelabel] {$1$} (v2) -- node[timelabel] {$3$}  (v3) -- node[timelabel] {$5$} (v4) -- node[timelabel] {$7$} (v5);
	\draw (v1) to [out=80,in=100,distance=2cm] node[timelabel] {$9$} (v3);

@@

\begin{proof}
    Let $D$ be an input matrix for problem \deltaExact\ of dimension $n \times n$.
    Let us fix the vertices of the corresponding graph $G$ of $D$ as $v_1, v_2, \dots, v_n$, where vertex $v_i$ corresponds to the row and column $i$ of matrix $D$.
    This can be done in polynomial time as we need to loop through the matrix $D$ once and connect vertices $v_i, v_j$ for which $D_{i,j} = 1$. At the same time we also check if $D_{i,i} = 0$, for all $i \in [n]$.
    When $G$ is constructed we run DFS algorithm on it and check if it has no cycles.
    If at any step we encounter a problem, our algorithm stops and returns a negative answer.

@@

\section{\deltaExact\ is NP-hard}

\begin{theorem}
	\deltaExactLong\ is \NP-hard.
\end{theorem}

\begin{proof}
	We present a polynomial-time reduction from the NP-hard problem 3-SAT~\cite{Karp1972Reducibility}. Here, we are given a formula $\phi$ in conjunctive normal form, where each clause contains exactly 3 literals (with three distinct variables). We construct an instance $D^\phi$ of \deltaExact\ as follows.

 We start by describing the vertex set of the underlying graph $G$ of  $D^\phi$.
\begin{itemize}
\item For each variable $x$ in $\phi$, we create three variable vertices $x, x^T, x^F$.
\item For each clause $c$ in $\phi$, we create one clause vertex $c$.
\item We add one additional super vertex $v$.
\end{itemize}
Next, we describe the edge set of $G$.
\begin{itemize}
\item For each variable $x$ in $\phi$ we add the following five edges: 
$x x^T, x x^F, x^T x^F, x^T v$ and $ x^Fv$.
\item For each pair of variables $x,y$ in $\phi$ with $x\neq y$ we add the following four edges: 
$x^T y^T, x^Ty^F, x^Fy^T$ and $x^Fy^F$.
\item For each clause $c$ in $\phi$ we add one edge for each literal. Let $x$ appear in $c$. If $x$ appears non-negated in $c$ we add edge $c x^T$. 
If $x$ appears negated in $c$ we add edge $c x^F$.
\end{itemize}
This finishes the construction of $G$.
For an example see~\cref{fig:NP-example}.

Now we specify the distances in the matrix $D^\phi$ between all vertex pairs.
For the sake of simplicity we write $D^\phi _{x,y}$ as $d(x,y)$,
where $x,y$ are two vertices. 
We start by setting the value of $d(x,y) = 1$ where $x$ and $y$ are two adjacent vertices in $G$.
\begin{itemize}
    %\item For each variable $x$ in $\phi$ we specify the following distances between the non-adjacent variable vertices:
    %$d(x_1,x_2)=2$. 
    %\item For pair of variable $x,y$ in $\phi$ with $x\neq y$ we specify the following 
    %$d(x_1,y^T)=d(x_1,y^F)=3$, $d(x_2,y^T)=d(x_2,y^F)=2$.
    \item For each variable $x$ in $\phi$ we specify the following distance to the super vertex $v$:
    $d(x,v)=2$. %, $d(x_2,v)=3$.

    \item For each clause $c$ in $\phi$ we specify the following distance to the super vertex $v$:
    $d(c,v)=2$
    \item Let $x$ be a variable that appears in clause $c$, then  we specify the following distance:
    $d(c,x)=2$.
    If $x$ appears non-negated in $c$ we specify the following distance:
    $d(c,x^F)=2$.
    If $x$ appears negated in $c$ we specify the following distance:
    $d(c,x^T)=2$.
    \item Let $x$ be a variable that does \emph{not} appear in clause $c$, then we specify the following distance:
    $d(c,x^T)=d(c,x^F)=2$.
\end{itemize}
All distances between non-adjacent vertex pairs that are not defined above are set to $\infty$.

This finishes the construction of an instance $D^\phi$ of \deltaExact\, which can clearly be done in polynomial time. In the remainder we show that $D^\phi$ is a yes-instance if and only if $\phi$ is satisfiable.

$(\Rightarrow)$: Assume the constructed instance $D^\phi$ of \deltaExact\ is a yes-instance. 
Then there exist a label $\lambda(e)$ for each edge $e\in E(G)$ such that for each vertex pair $u,w$ in the temporal graph $(G,\lambda)$ we have that a fastest temporal path from $u$ to $w$ is of duration $d(u,w)$. 
%In particular, for any two vertices $u,w$ with $d(u,w)=\infty$ we have that there does not exist a temporal path from $u$ to $w$ in $(G,\lambda)$.

We construct a satisfying assignment for $\phi$ as follows. For each variable $x$, if $\lambda(x x^T)=\lambda(x^T v)$, then we set $x$ to \textsc{True}, otherwise we set $x$ to \textsc{False}.

@@

Let $\lambda(x x^T)=t$ and $\lambda(x x^F)=t'$, for an arbitrary variable $x$. 
If both $\lambda(x^T v)\neq t+1$ and $\lambda(x^F v)\neq t'+1$, then $d(x,v)>2$, which is a contradiction. 
Thus, for every variable $x$, we have that $\lambda(x^T v)= t+1$ or $\lambda(x^F v)= t'+1$ (or both). 
In particular, this means that if $\lambda(x x^F)=\lambda(x^F v)$, then we set $x$ to \textsc{False}, since in this case $\lambda(x x^T)\neq\lambda(x^T v)$.

@@

Now assume for contradiction that the described assignment is not satisfying. Then there exists a clause $c$ that is not satisfied. Recall that we require $d(c,v)=2$. Hence, we must have a temporal path consisting of two edges from $c$ to $v$ such that the two edges have consecutive labels. By construction of $G$ there are three candidates for such a path, one for each literal of $c$. 
Assume w.l.o.g\ that $x$ appears in $c$ non-negated (the case of a negated appearance of $x$ is symmetrical) and that the temporal path realizing $d(c,v)=2$ goes through vertex $x^T$. 
Then we have that $\lambda(c x^T)=\lambda(x^T v)-1$. Furthermore, since $d(c,x)=2$ we also have that $\lambda(c x^T)=\lambda(x x^T)-1$. It follows that $\lambda(x x^T)=\lambda(x^T v)$. However, this implies that $x$ is set to \textsc{True} in the satisfying assignment and thus the clause $c$ is satisfied, a contradiction.

@@

We construct a labeling $\lambda$ as follows.
\begin{itemize}
    \item All edges incident with a clause vertex $c$ obtain label one.
    %\item For each variable $x$, we set $\lambda(\{x^T,x^F\})=1$.
    \item If variable $x$ is set to \textsc{True}, we set $\lambda(x^F v)=3$.
    \item If variable $x$ is set to \textsc{False}, we set $\lambda(x^T v)=3$.
    \item We set the labels of all other edges to two.
\end{itemize}

@@

    If $x$ is set to \textsc{True}, then there is a temporal path from $x$ to $v$ via $x^F$ such that $\lambda(x x^F)=2$ and $\lambda(x^F v)=3$. If $x$ is set to \textsc{False}, then there is a temporal path from $x$ to $v$ via $x^T$ such that $\lambda(x x^T)=2$ and $\lambda(x^T v)=3$.

@@

    Since we have a satisfying assignment there is a variable $x$ appearing in $c$ that is set to a truth-value that satisfies $c$. If $x$ appears non-negated in $c$ (and hence is set to \textsc{True}), then there is a temporal path from $c$ to $v$ through $x^T$ such that $\lambda(c x^T)=1$ and $\lambda(x^T v)=2$. If $x$ appears negated in $c$ (and hence is set to \textsc{False}), then there is a temporal path from $c$ to $v$ through $x^F$ such that $\lambda(c x^F)=1$ and $\lambda(x^F v)=2$.
    \item Let $x$ be a variable that appears in clause $c$.
    If $x$ appears non-negated in $c$ we have $d(c,x)=d(c,x_F)=2$:

    There is a temporal path from $c$ to $x$ via $x^T$ and also a temporal path from $c$ to $x^F$ via $x^T$ such that $\lambda(c x^T)=1$ and $\lambda(x x^T)=\lambda(x^T x^F)=2$.

@@

    There is a temporal path from $c$ to $x$ via $x^F$ and also a temporal path from $c$ to $x^T$ via $x^F$ such that $\lambda(c x^F)=1$ and $\lambda(x x^F)=\lambda(x^T x^F)=2$.
    \item Let $x$ be a variable that does \emph{not} appear in clause $c$, then we have $d(c,x^T)=d(c,x^F)=2$:

    Let $y$ be a variable that appears non-negated in $c$ (the case where $y$ appears negated is symmetrical). Then there is a temporal path from $c$ to $x^T$ via $y^T$ and also a temporal path from $c$ to $x^F$ via $y^T$ such that $\lambda(c y^T)=1$ and $\lambda(y^T x^T)=\lambda(y^T x^F)=2$.
\end{itemize}
Lastly, we show that all non-adjacent vertex pairs $u,w$ with $d(u,w)=\infty$ are not temporally connected in $(G,\lambda)$.
\begin{itemize}
\item For all pairs of clause vertices $c,c'$ we have $d(c,c')=\infty$: 
\todo[inline]{New matrix entries D for period Delta: d(c,c')=Delta+1 (using three edges, with labels L, L+1, L).}

@@

All edges incident with $v$ have labels two or three. All edges incident with $x$ have label two and all edges incident with $c$ have label one. We immediately have that there is no temporal path from $v$ to $c$. Furthermore, since $x$ and $v$ are non-adjacent, we also have that there is no temporal path from $v$ to $x$.
\end{itemize}
%
\begin{figure}[htb]
	\noindent
	\makebox[\textwidth]{
		\centering
\begin{tikzpicture}%[xscale=2]

@@

    Given an instance $H$ of \textsc{Multicolored Clique}, we construct an instance $D$ of \deltaExactLong\ as follows. To ease the presentation, we first describe the underlying graph $G$ that is implicitly defined by the entries $D_{v,v'}=1$, that is, the pairs of vertices $v,v'$ that should be connected by a temporal path of duration one, meaning that there needs to an edge connecting the two vertices. Afterwards, we describe the remaining entries of $D$.
    We will construct $G$ using several gadgets.

    \subparagraph{Edge Selection Gadget.} We first introduce an \emph{edge selection gadget for color combination $i,j$} with $i<j$. We start with describing the vertex set of the gadget.
    \begin{itemize}
        \item A set $X_{i,j}$ of vertices $x_1, x_2, \ldots, x_m$.
        %\item A set $Y_{i,j}$ of vertices $y_1, y_2, \ldots, y_m$.
        %\item A set $Z_{i,j}$ of vertices $z_1, z_2, \ldots, z_m$.
        \item Vertex sets $U_1, U_2, \ldots, U_m$ with $U_\ell=\{u^\ell_0, u^\ell_1, u^\ell_2,\ldots, u^\ell_n, u^\ell_{n+1}, u^\ell_{n+2}, u^\ell_{2n}\}$ for all $\ell\in[m]$. Let $U_{i,j}=\bigcup_{\ell\in[n]} U_\ell$.
        \item Two special vertices $v_{i,j}^\star,v_{i,j}^{\star\star}$.
    \end{itemize}
    The gadget has the following edges.
    \begin{itemize}
        \item For all $\ell\in [m]$ we have edge $\{x_\ell,v_{i,j}^\star\}$, $\{v_{i,j}^\star,u^\ell_0\}$, and $\{u^\ell_{2n},v_{i,j}^{\star\star}\}$.

        %Furthermore, we have edges $\{y_\ell,u^\ell_1\}$ and $\{z_\ell,u^\ell_{2n}\}$.
        \item For all $\ell\in [m]$ and $\ell'\in [2n]$, we have edge $\{u^\ell_{\ell'-1},u^\ell_{\ell'}\}$.
    \end{itemize}
    %This finishes the construction of the edge selection gadget.

    \subparagraph{Verification Gadget.} For each color $i$, we introduce the following vertices. What we describe in the following will be used as a \emph{verification gadget for color $i$}.
    \begin{itemize}
        \item We have one vertex $y^i$, one vertex $v^i_0$, and $k-1$ vertices $v^i_\ell$ for $\ell\in[k]\setminus\{i\}$.
        \item For every $\ell\in[m]$ and every $j\in[k]\setminus\{i\}$ we have $n$ vertices $a^{i,j,\ell}_1,a^{i,j,\ell}_2,\ldots,a^{i,j,\ell}_{n}$ and $n$ vertices $b^{i,j,\ell}_1,b^{i,j,\ell}_2,\ldots,b^{i,j,\ell}_{n}$.
        \item If $1\neq i\neq k$, then we have a set $\hat{U}_i$ of $n+3$ additional vertices $\hat{u}^i_1,\hat{u}^i_2,\ldots,\hat{u}^i_{n+3}$.
    \end{itemize}
    We add the following edges. We add edge $\{y^i,v^i_0\}$.
    %For every $j\in[k]\setminus\{i,i+1\}$ and every $\ell\in[m]$ we add edge $\{v^i_{j-1},\hat{u}^{i,j,\ell}_1\}$. For every $\ell\in[m]$ we add edge $\{\hat{u}^i_{n+3},\hat{u}^{i,i+1,\ell}_1\}$.
    For every $\ell\in[m]$, every $j\in[k]\setminus\{i\}$, and every $\ell'\in[n-1]$ we add edge $\{a^{i,j,\ell}_{\ell'},a^{i,j,\ell}_{\ell'+1}\}$ and we add edge $\{b^{i,j,\ell}_{\ell'},b^{i,j,\ell}_{\ell'+1}\}$.
    %Assume $1\neq i \neq k$. We describe the two cases $i=1$ and $i=k$ afterwards. 
    %If $i\neq 1$, let $e_\ell\in F_{1,i}$ let $w^i_{\ell'}\in W_i$ be incident with $e_\ell$. Then we add edge $\{v^i_0,u^\ell_{\ell'}\}$ between $v^i_0$ and the vertex $u^\ell_{\ell'}$ of the edge selection gadget of color combination $1,i$.

    Let $1\le j<i$ (skip if $i=1$), let $e_\ell^{j,i}\in F_{j,i}$, and let $w^i_{\ell'}\in W_i$ be incident with $e_\ell^{j,i}$. Then we add edge $\{v_{j-1}^i,a^{i,j,\ell}_{1}\}$ and we add edge $\{a^{i,j,\ell}_{n},u^\ell_{\ell'-1}\}$ between $a^{i,j,\ell}_{n}$ and the vertex $u^\ell_{\ell'-1}$ of the edge selection gadget of color combination $j,i$.
    Furthermore, we add edge $\{v_{j}^i,b^{i,j,\ell}_{1}\}$ and edge $\{b^{i,j,\ell}_{n},u^\ell_{\ell'}\}$ between $b^{i,j,\ell}_{n}$ and the vertex $u^\ell_{\ell'}$ of the edge selection gadget of color combination $j,i$.

    If $1\neq i\neq k$, then we add edge $\{v^i_{i-1},\hat{u}^i_1\}$ and for all $\ell\in[n+2]$ we add edge $\{\hat{u}^i_\ell,\hat{u}^i_{\ell+1}\}$. 
    Let $e_\ell^{i,i+1}\in F_{i,i+1}$ and let $w^i_{\ell'}\in W_i$ be incident with $e_\ell^{i,i+1}$. If $i\neq 1\neq k$, then we add edge $\{\hat{u}^i_{n+3},a^{i,i+1,\ell}_{1}\}$ and edge $\{a^{i,i+1,\ell}_{n},u^\ell_{n+\ell'-1}\}$ between $a^{i,i+1,\ell}_{n}$ and the vertex $u^\ell_{n+\ell'-1}$ of the edge selection gadget of color combination $i,i+1$. If $i=1$, then we add edge $\{v^i_0,a^{i,i+1,\ell}_{1}\}$ and edge $\{a^{i,i+1,\ell}_{n},u^\ell_{n+\ell'-1}\}$ between $a^{i,i+1,\ell}_{n}$ and the vertex $u^\ell_{n+\ell'-1}$ of the edge selection gadget of color combination $i,i+1$.
Furthermore, we add edge $\{v_{i+1}^i,b^{i,i+1,\ell}_{1}\}$ and edge $\{b^{i,i+1,\ell}_{n},u^\ell_{\ell'}\}$ between $b^{i,i+1,\ell}_{n}$ and the vertex $u^\ell_{\ell'}$ of the edge selection gadget of color combination $i,i+1$.

    Let $i+1<j\le k$ (skip if $i=k$), let $e_\ell^{i,j}\in F_{i,j}$, and let $w^i_{\ell'}\in W_i$ be incident with $e_\ell^{i,j}$. Then we add edge $\{v_{j-1}^i,a^{i,j,\ell}_{1}\}$ and edge $\{a^{i,j,\ell}_{n},u^\ell_{n+\ell'-1}\}$ between $a^{i,j,\ell}_{n}$ and the vertex $u^\ell_{n+\ell'-1}$ of the edge selection gadget of color combination $i,j$.
    Furthermore, we add edge $\{v_{j}^i,b^{i,j,\ell}_{1}\}$ and edge $\{b^{i,j,\ell}_{n},u^\ell_{n+\ell'}\}$ between $b^{i,j,\ell}_{n}$ and the vertex $u^\ell_{n+\ell'}$ of the edge selection gadget of color combination $i,j$.
